Decoupled Fallback: A Portable Single-Pass GPU Scan

### 2025-02-28 paper.tex
making fig 1 caption accurate

\subsection{Why is Reliance on Forward-Progress Guarantees a Portability Problem?}
The absence of FPG in key hardware segments poses a critical challenge for developers. Our testing (Figure~\ref{fig:dist}) shows that, at best, running \emph{Chained Scan} without FPG results in performance worse than \emph{Reduce-then-Scan}, and frequently triggers \emph{Timeout Device Recovery}\footnote{Operating systems handle unresponsive programs differently. On Windows, the Timeout Detection and Recovery (TDR) system resets the GPU after approximately 2 seconds. macOS has a timeout watchdog that, interestingly, does not terminate a spinning kernel---allowing us to obtain these results. On Linux, behavior varies by GPU driver.} (TDR), crashing the program.
\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{graphics/dist.pdf}
  \caption{Execution times of \emph{Decoupled Lookback} on an FPG-lacking device, Apple M1 Max. The TDR threshold represents the duration after which the OS, assuming a non-terminating kernel, kills the driver.\label{fig:dist}}
\end{figure}
Exacerbating this issue, no major graphics API provides a mechanism to query FPG support. This lack of transparency forces developers to design algorithms without knowing whether critical guarantees are available on the target hardware. In shader environments, where cross-platform portability is essential, this uncertainty compels a fallback to older, less efficient strategies like \emph{Reduce-then-Scan}, sacrificing performance for broader compatibility.

>>>>

\subsection{Why is Reliance on Forward-Progress Guarantees a Portability Problem?}
The absence of FPG in key hardware segments poses a critical challenge for developers. Our testing (Figure~\ref{fig:dist}) shows that, at best, running \emph{Chained Scan} without FPG results in performance worse than \emph{Reduce-then-Scan}, and frequently triggers \emph{Timeout Device Recovery}\label{sec:tdr}\footnote{Operating systems handle unresponsive programs differently. On Windows, the Timeout Detection and Recovery (TDR) system resets the GPU after approximately 2 seconds. macOS has a timeout watchdog that, interestingly, does not terminate a spinning kernel---allowing us to obtain these results. On Linux, behavior varies by GPU driver.} (TDR), crashing the program.
\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{graphics/dist.pdf}
  \caption{Execution times of \emph{Decoupled Lookback} on an FPG-lacking device, Apple M1 Max. The TDR threshold here is Windows's threshold for resetting the GPU\@; as noted in Section~\ref{sec:tdr}, MacOS handles this differently.\label{fig:dist}}
\end{figure}
Exacerbating this issue, no major graphics API provides a mechanism to query FPG support. This lack of transparency forces developers to design algorithms without knowing whether critical guarantees are available on the target hardware. In shader environments, where cross-platform portability is essential, this uncertainty compels a fallback to older, less efficient strategies like \emph{Reduce-then-Scan}, sacrificing performance for broader compatibility.
